---
title: "Data-Clean-Up-Invert-Riparian"
author: "Jonathan Ray, Callie Wallace, Chelsey Burks"
date: "2023-02-27"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
#loading packages for tidyverse and ggplot
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(readr)

```


```{r datasets Jonathan}
#recalling individual data sets from Jonathan's D08 MAYF
data11 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20120.001.inv_taxonomyProcessed.2017-07.expanded.20221122T184910Z.csv")
data12 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20120.001.inv_taxonomyProcessed.2018-07.expanded.20221128T184824Z.csv")
data13 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20120.001.inv_taxonomyProcessed.2019-07.expanded.20221128T173134Z.csv")
data14 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20120.001.inv_taxonomyProcessed.2020-07.expanded.20221128T212351Z.csv")
data15 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20120.001.inv_taxonomyProcessed.2021-07.expanded.20221129T232348Z.csv")
data16 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20191.001.rip_percentComposition.2017-08.basic.20230112T002045Z.csv")
data17 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20191.001.rip_percentComposition.2018-08.basic.20230112T002023Z.csv")
data18 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20191.001.rip_percentComposition.2019-08.basic.20230112T001825Z.csv")
data19 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20191.001.rip_percentComposition.2021-08.basic.20230112T002158Z.csv")
data110 <- read.csv("../data/processed/D08 MAYF/NEON.D08.MAYF.DP1.20191.001.rip_percentComposition.2022-08.basic.20230112T002125Z.csv")

```

```{r dataset Callie}
#recalling individual data sets from Callie's D17 BIGC
data21 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20120.001.inv_taxonomyProcessed.2018-07.expanded.20221128T191616Z.csv")
data22 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20120.001.inv_taxonomyProcessed.2019-07.expanded.20221128T162452Z 2.csv")
data23 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20120.001.inv_taxonomyProcessed.2021-07.expanded.20221129T222632Z.csv")
data24 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20191.001.rip_percentComposition.2018-07.basic.20230112T002002Z.csv")
data25 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20191.001.rip_percentComposition.2019-07.basic.20230112T002213Z.csv")
data26 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20191.001.rip_percentComposition.2021-08.basic.20230112T001807Z.csv")
data27 <- read.csv("../data/processed/D17 BIGC/NEON.D17.BIGC.DP1.20191.001.rip_percentComposition.2021-08.basic.20230112T001807Z.csv")
```

```{r datasets Chelsey}
#recalling individual data sets from Chelsey's D18 OKSR
data31 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20120.001.inv_taxonomyProcessed.2017-07.expanded.20221122T171622Z.csv")
data32 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20120.001.inv_taxonomyProcessed.2018-07.expanded.20221128T182544Z.csv")
data33 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20120.001.inv_taxonomyProcessed.2019-07.expanded.20221128T174721Z.csv")
data34 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20120.001.inv_taxonomyProcessed.2020-07.expanded.20221128T220306Z.csv")
data35 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20120.001.inv_taxonomyProcessed.2021-07.expanded.20221129T222833Z.csv")
data36 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20191.001.rip_percentComposition.2017-07.basic.20230112T002307Z.csv")
data37 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20191.001.rip_percentComposition.2019-07.basic.20230112T001757Z.csv")
data38 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20191.001.rip_percentComposition.2021-07.basic.20230112T002024Z.csv")
data39 <- read.csv("../data/processed/D18 OKSR/NEON.D18.OKSR.DP1.20191.001.rip_percentComposition.2022-07.basic.20230112T001943Z.csv")

```

```{r combined}
all_macroinvert <- bind_rows(data11,data12,data13,data14,data15,data21,data22,data23,data31,data32,data33,data34,data35) #combining all macroinvertebrate data into one data frame for further merging
all_macroinvert_year <- all_macroinvert %>% 
  mutate(year = floor_date(as.Date(collectDate), unit = "year")) %>% #Collapsing collectDate into a year column in order to merge by date
  group_by(siteID, year) %>% 
  select("siteID", "genus", "year") #removing unneeded columns
all_riparian <- bind_rows(data16,data17,data18,data19,data110,data24,data25,data26,data27,data36,data37,data38,data39) #combining all riparian data into one data frame for further merging
all_riparian_year <- all_riparian %>% 
  mutate(year = floor_date(as.Date(startDate), unit = "year")) %>% #Same as line 65
  group_by(siteID, year) %>% 
  select("siteID", "canopyCoverPercent", "year") #Same as line 67
combined_data <- left_join(all_macroinvert_year,all_riparian_year, by = c("siteID","year")) %>% #Joining macroinvertebrate data and riparian data into a single data frame based on shared siteID and year for analysis
  relocate(year, .before = siteID) #moving year column before siteID to improve aesthetic

```

```{r histogram}
combined_data %>% 
  ggplot(aes(x=canopyCoverPercent)) +
  geom_histogram() #ggplot histogram of count data for canopy cover percent to determine distribution of variable
combined_data %>% 
  group_by(siteID,year) %>% 
  summarise(richness = unique(length(genus))) %>% 
  ggplot(aes(x = richness)) +
  geom_histogram() #ggplot histogram of count data for species richness to examine distribution of variable
```

```{r grouping}
combined_data_group <- combined_data %>% 
  group_by(siteID, year) %>% 
  summarise(species.richness = unique(length(genus)), cover.average = mean(canopyCoverPercent)) #creating species richness and average canopy cover per site per year. Since species richness is measured for the entire year, average canopy cover must be calculated as well.

```

```{r visualization}
combined_data_group %>% 
  ggplot(aes(x = cover.average, y = species.richness)) +
  geom_point(aes(color = combined_data_group$siteID)) + #adding in data points to scatter
  geom_smooth(method = "lm", color = "black") + #adding linear regression line to display our test 
  xlab("Average Percent Canopy Cover") +
  ylab("Species Richness") +
  guides(color = guide_legend(title = "Site"))
  
```
Figure 1: This visualization shows a comparison of species richness (y-axis) and canopy coverage (x-axis). The three sites shown ,in differing colors, compare the coverage and the species richness by year. The pattern shows that those sites with more canopy cover have a higher richness in aquatic macro-invertebrates. 

```{r second vis}
combined_data_group$year_character <- as.character(combined_data_group$year) #changing variable type to character to improve aesthetic of bar color
combined_data_group %>%
  drop_na() %>% #removing years with no recordings to remove extra spaces
  group_by(siteID, year_character) %>% 
  ggplot(aes(x = siteID, y = species.richness, fill = year_character)) + 
  geom_col(position = "dodge") + #bar plot of species richness grouped by site and year to make sure they are similar across years for the same site
  xlab("Site") +
  ylab("Species Richness")

```
Figure 2: On the x-axis of this figure, three sites are labeled and colored by year, and the y-axis is aquatic macro-invertebrates richness. Each color signifies the year the data was collected. The pattern shows that in Upper Big Creek (BIGC) there was an increase in richness, Mayfield Creek had a decrease in richness, and Oksrukuyik Creek had no change in richness.  

```{r stats}
model_test <- lm(species.richness ~ cover.average, data = combined_data_group) #establishing linear regression model test for species.richness predicted by cover.average
summary(model_test) #print out of summary for analysis
plot(model_test) #plots for model assumptions

```
Our Null hypothesis is that percent cover has no true effect on species richness of macroinvertebrate species. The P-value is 0.296, which is much greater than our alpha value of 0.05, so we fail to reject our null hypothesis. The R squared value is also incredibly small, indicating there is very little correlation between our average percent coverage and species richness of aquatic macroinvertebrates. Based on our plots, some assumptions appear to be violated. Examing the Q-Q plot, residuals do not appear to be normally distributed while the residual plot appears to show a roughly linear relationship between species richness and average canopy cover. Finally, the scale-location plot shows that the variance of the model residuals were not the same for all values. 

```{r third vis}
combined_data_group$year_character <- as.character(combined_data_group$year) #same as line 102
combined_data_group %>%
  drop_na() %>% #same as 104
  group_by(siteID, year_character) %>% 
  ggplot(aes(x = siteID, y = cover.average, fill = year_character)) +
  geom_col(position = "dodge") + #bar plot of cover.average grouped by siteID and year to make sure average canopy cover was similar across sampled years for each site
  xlab("Site") +
  ylab("Average Percent Canopy Cover")
```
Figure 3: A bar graph of average canopy cover (y-axis) grouped by site (x-axis) and year (bar color). Higher bars represent large amounts of average canopy cover where the Alaska site (OKSR) had very little to no tree cover. Not all sites had a 2017 or a 2021 recording, but all three sites had similar average percent canopy cover across all three years. 